### 2022-06-28 manuscript.tex
Merge pull request #246 from SORTEE-Github-Hackathon/RCO-discussion-conclusion [ci skip]
This build is based on
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/cc849099b23d210f6928264fd6a353fdcc54862d.

This commit was created by the following CI build and job:
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/cc849099b23d210f6928264fd6a353fdcc54862d/checks
https://github.com/SORTEE-Github-Hackathon/manuscript/actions/runs/2578285311

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/773da7b1a68108cf2ee8aa5b261b6f4ab75fb2ab/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/773da7b1a68108cf2ee8aa5b261b6f4ab75fb2ab}{SORTEE-Github-Hackathon/manuscript@773da7b}
on June 28, 2022.

@@

There have been many calls for researchers outside of the software development community to join the 73 million GitHub users for their collaborative research\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-1Du6fzB8g}{38},\protect\hyperlink{ref-UsTxAq4f}{52}}.
This call comes in light of both the continual shift toward open-science and increasing computational and data requirements in EEB.
Until now, resources and practical guidance specifically focused on using GitHub within the EEB community have been dispersed in blog posts and video tutorials (\protect\hyperlink{tips}{Box 2} Box\textsuperscript{\protect\hyperlink{ref-box:tips}{\textbf{box:tips?}}}).
We felt these resources have been extremely useful for us to learn to use GitHub in our own work, and that a collation of the main ideas into one medium, while adding on our personal perspectives, would be of use to the EEB community.

In this paper, we described 12 tractable ways that EEB researchers can leverage GitHub to enable more transparent and collaborative research (Figure \ref{fig:scatterblob}).
Many of the examples are specifically meant for first-time GitHub users and can likely be adopted with just a little practice (storing data, creating virtual notebooks, making code citable).
For example, storing code and data and making it citable generally just involve creating a repository on GitHub, pushing code to the repository, and then going through the necessary steps (e.g., connecting a GitHub repository with Zenodo; see below) for creating a DOI for the repository.
These actions are often covered in any introductory tutorial on GitHub and take little effort to implement.
On the other hand, some other examples we described here, including material development, web hosting, and delivery, may require a greater time commitment, but have the potential to make EEB research more open, accessible, and collaborative than ever before.
Managing full research projects or research labs on GitHub will require careful thought as to how to delegate tasks such as reviewing pull requests or creating issues, as well as thought as to how modular to make the research project or research lab (i.e., which repository will be used for what, and how many repositories are needed).
For example, collaboratively writing a paper using GitHub, as we have done here, will involve a learning curve for co-authors less familiar with the intricacies of GitHub, and also require overhead to set up the repository using GitHub actions.
Despite the many potential applications of GitHub to EEB research, we acknowledge that there will still be many times when researchers might look to other platforms for research collaboration.

@@

Though we see GitHub as a useful tool for collaboration in EEB, we describe two use cases where, to our knowledge, GitHub's features still fall short of the type of highly collaborative work emblematic of EEB research.
First, real-time collaborative editing (e.g., as on a shared Google Doc or a Word document stored on Dropbox) is not possible on GitHub.
There are websites outside of the GitHub ecosystem that are built on top of the GitHub architecture that allow real-time collaborative editing (e.g., hackMD (\url{https://hackmd.io/}) or replit (\url{https://replit.com/})).
We used HackMD at two key points in writing our manuscript when real-time co-writing was essential: when taking meeting notes and writing the outline of our paper.
Second, we looked to other software when working on figures and tables.
Though creating tables and figures on GitHub using markdown or other scripting languages is possible, we found that it was not practical at the early brainstorming stages.
We needed to rapidly iterate on figure and table design, share feedback through comments, and merge/reorder ideas when necessary.
For these reasons, we used Google Slides for working on figures and Google Sheets for working on tables.
As our figures and tables moved towards more finalized forms, some co-authors chose to create the tables and figures using R and Markdown which could then be tracked using the same version control system as the rest of manuscript.

@@

First, there may be a hesitation to use GitHub due to the somewhat steep learning curve to using the platform combined with limited instruction available through traditional university courses.
When GitHub is taught within an EEB context, it is usually accompanying coursework in topics such as statistical programming, and some students may find it overwhelming to juggle learning git alongside scripting languages, statistical theory, and file system navigation, especially when many may also be new to using command-line interfaces in general.
Instructors likewise may confuse the expected digital literacy of younger students with computational fluency, even when modern technology increasingly abstracts many relevant concepts through search optimization and preponderant IDEs (Integrated Development Environments), or `point-and-click' user interfaces.

Second, while many EEB researchers take advantage of GitHub for individual use, collaborative use may lag due to how researchers traditionally divide labor within projects.
Despite broad utility, GitHub remains a tool predominantly used by computer scientists and software developers.
EEB researchers may take the view that GitHub is a platform that only needs to be used by individuals writing code, and may silo those aspects of projects to a single individual.
Those assumptions may obscure the utility of GitHub for tasks other than traditional data analysis and code development, or how GitHub can facilitate the integration of code with non-coding aspects of projects through the practice of repository design.
However, we emphasize that there are opportunities for collaboration using GitHub by researchers of all skill levels or time constraints (Table \ref{tbl:roles}); for example, project stakeholders can provide a list of use-cases or highlight important conceptual components of a project using the issues or discussions functionality of GitHub.

A third barrier may come from general reluctance to share data and code publicly, or technical and logistical issues.
The default public nature of GitHub usage can add additional pressure to students and scientists learning to use the platform.
We also note that large file storage is discouraged (and limited) on GitHub but add-ons do exist (e.g.~\url{https://git-lfs.github.com/}) that permit data storage, and increasing integration between platforms (e.g.\textsuperscript{\protect\hyperlink{ref-lx49NGto}{53}}) allows data to be stored away from GitHub and linked dynamically.
We suspect a major additional barrier to EEB researchers is a distinct lack of GitHub help documents for non-English researchers in ecology and evolution, meaning that EEB researchers potentially miss the opportunity to fully understand the importance of version control, reproducibility, and other benefits of GitHub.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{1.0000}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
10 Tips for getting started in GitHub
\end{minipage} \\
\midrule
\endhead
1. \textbf{Check for an existing solution to your problem.} The GitHub Help webpage (\url{https://docs.github.com/en}) contains extensive and detailed documents with helpful screenshots. It is a good starting point for handling an issue, and has troubleshooting tips for specific problems. Alternatively, consider Tweeting your issue. There is a large community of GitHub users around the world who have likely faced analogous problems and may be able to provide quick solutions. Third, try to follow blogs (e.g., \url{https://github.blog}), Twitter accounts or YouTube channels that regularly post practical solutions about the most widely-used web platform for common GitHub issues. \\
2. \textbf{Consider taking free courses.} such as those from Software Carpentry\textsuperscript{\protect\hyperlink{ref-pjy75gHr}{54}} and sharing these courses with your lab members or colleagues. \\
3. \textbf{Take advantage of GitHub as an asynchronous working tool for team-based projects.} See the repository for this paper (\url{https://github.com/SORTEE-Github-Hackathon/manuscript}) as an example of a collaborative manuscript that includes discussions, issues, and a website via GitHub. \\
4. \textbf{The GitHub Learning Lab (\url{https://lab.github.com}).} allows you to learn GitHub basics through short projects and tasks, and allows you to get feedback from their Learning Lab bot. \\
5. \textbf{Check out the following markdown cheatsheet (\url{http://markdownguide.org/basic-syntax}).} so that you can write clear metadata README files for your repositories. \\
6. \textbf{The Jenny Bryan universe of GitHub material.} provides a thorough and accessible introduction for a multitude of research-related uses for GitHub, and includes a book\textsuperscript{\protect\hyperlink{ref-ZvrOcg9w}{55}}, statistics course\textsuperscript{\protect\hyperlink{ref-6CMMeSeD}{56}} and academic article\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}. \\
7. \textbf{Don't be afraid of trial and error.} One of the best ways to learn GitHub is the trial and error method. Learning from your own mistakes can be the better way to master your GitHub abilities. In any case, GitHub has the advantage of making it easy to go back to any steps that you desire via version controlling if you make mistakes. \\
8. \textbf{If you are an educator, include lectures on reproducibility and tools for creating reproducible workflows in the curricula.} Some graduate programs include coursework on course Rmarkdown and GitHub. Getting students started with these tools earlier will prevent the resistance that comes from working with a less reproducible workflow for a longer period of time. (see example \url{https://github.com/rmcelreath/stat_rethinking_2022}) \\
9. \textbf{Try to begin committing with GUI (Graphical user interface) tools.} like GitHub Desktop (\url{https://desktop.github.com}), git-gui (\url{https://git-scm.com/docs/git-gui}), RStudio (\url{https://www.rstudio.com}), Visual Studio Code (\url{https://code.visualstudio.com}), Atom (\url{https://atom.io}), GitKraken (\url{https://www.gitkraken.com}) tools instead of CLI (Command line interface) tools such as Terminal or Console for more advanced features. \\
10. \textbf{Get help deciphering GitHub Notifications.} Try using tools like Octobox (\url{https://octobox.io}) to disentangle and manage multiple notifications from distinct GitHub projects. \\
\bottomrule
\end{longtable}

@@

We provide 12 practical ways that Ecologists and Evolutionary Biologists can incorporate GitHub into their research workflows, and include definitions (\protect\hyperlink{definitions}{Box 1}) and key user groups (Figure \ref{fig:github-diagram}) that can help researchers prioritize which GitHub skills to learn first.
Some GitHub uses are highly collaborative (e.g., open science discussion and collaborative code editing) while others are focused on individual actions (e.g., storing code/data, building a website).
Regardless of the degree of collaboration, GitHub use in ecology and evolution has the potential to make the field more open and transparent than ever before.
Our paper provides the most comprehensive review of how EEB researchers can use GitHub to date, and we encourage EEB researchers at any career stage studying any topic to try GitHub as a platform for sharing and collaboration.

@@

\leavevmode\vadjust pre{\hypertarget{ref-lx49NGto}{}}%
\CSLLeftMargin{53. }
\CSLRightInline{Connect GitHub to a Project - OSF Support. \url{https://help.osf.io/article/211-connect-github-to-a-project}.}

\leavevmode\vadjust pre{\hypertarget{ref-pjy75gHr}{}}%
\CSLLeftMargin{54. }
\CSLRightInline{Madicken Munk \emph{et al.} swcarpentry/git-novice: Software Carpentry: Version Control with Git, June 2019. (2019) doi:\href{https://doi.org/10.5281/zenodo.3264950}{10.5281/zenodo.3264950}.}

\leavevmode\vadjust pre{\hypertarget{ref-ZvrOcg9w}{}}%
\CSLLeftMargin{55. }
\CSLRightInline{Hester, J. B., the STAT 545 TAs, Jim. \emph{\href{https://happygitwithr.com/}{Let's Git started \textbar{} Happy Git and GitHub for the useR}}.}

\leavevmode\vadjust pre{\hypertarget{ref-6CMMeSeD}{}}%
\CSLLeftMargin{56. }
\CSLRightInline{Bryan, J. \& TAs, T. S. 545. \emph{\href{https://stat545.com/}{STAT 545}}.}

>>>>

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/cc849099b23d210f6928264fd6a353fdcc54862d/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/cc849099b23d210f6928264fd6a353fdcc54862d}{SORTEE-Github-Hackathon/manuscript@cc84909}
on June 28, 2022.

@@

There have been many calls for researchers outside of the software development community to join the 73 million GitHub users for their collaborative research\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-1Du6fzB8g}{38},\protect\hyperlink{ref-UsTxAq4f}{52}}.
This call comes in light of both the continual shift toward open-science and increasing computational and data requirements in EEB.
Until now, resources and practical guidance on using GitHub within the EEB community have been dispersed in blog posts and video tutorials (\protect\hyperlink{tips}{Box 2}).
These resources have been extremely useful for learning to use GitHub in our own work.
We also feel that a collation of the main uses of GitHub in EEB into one medium, while adding on our personal perspectives, will be useful to the EEB community.

In this paper, we described 12 tractable ways that EEB researchers can leverage GitHub to enable more transparent and collaborative research (Figure \ref{fig:scatterblob}).
Many of the examples are specifically meant for first-time GitHub users and can likely be adopted with just a little practice (storing data, creating virtual notebooks, making code citable).
For example, storing code and data and making it citable involve creating a repository on GitHub, pushing code to the repository, and then going through the necessary steps (\emph{e.g.}, connecting a GitHub repository with Zenodo; see below) for creating a DOI.
These actions are often covered in GitHub tutorials and take little effort to implement.
On the other hand, some examples we described here, including course material development, web hosting, and delivery, may require a greater time commitment, but have the potential to make EEB research more open, accessible, and collaborative than ever before.
Managing full research projects or research labs on GitHub will require careful thought as to how to delegate tasks such as reviewing pull requests or creating issues.
For example, collaboratively writing a paper using GitHub, as we have done here, may involve a learning curve for co-authors less familiar with the intricacies of GitHub, and requires overhead to set up the repository using GitHub actions.
Despite the many potential applications of GitHub to EEB research, we acknowledge that there will still be many times when researchers might look to other platforms for research collaboration.

@@

Though we see GitHub as a useful tool for collaboration in EEB, we describe two use cases where, to our knowledge, GitHub's features still fall short of the highly collaborative work emblematic of EEB research.
First, real-time collaborative editing (\emph{e.g.}, as on a shared Google Doc or a Word document stored on Dropbox) is not possible on GitHub.
There are websites outside of the GitHub ecosystem, built on top of the GitHub architecture, that allow real-time collaborative editing (\emph{e.g.}, hackMD (\url{https://hackmd.io/})).
We used HackMD at two key points in writing our manuscript when real-time co-writing was essential: when taking meeting notes and writing the outline of our paper.
Second, we looked to other software when working on figures and tables.
Though creating tables and figures on GitHub using markdown or other scripting languages is possible, we found that it was not practical at the early brainstorming stages.
We used Google Slides for working on figures and Google Sheets for working on tables.
As our figures and tables moved towards finalized forms, some co-authors chose to create the tables and figures using R and Markdown which were then tracked using the same version control system as the rest of manuscript.

@@

First, there may be a hesitation to use GitHub due to the somewhat steep learning curve to using the platform combined with limited instruction available through traditional university courses.
When GitHub is taught within an EEB context, it is usually accompanying coursework in topics such as statistical programming, and some students may find it overwhelming to juggle learning git alongside scripting languages, statistical theory, and file system navigation, especially when many may also be new to using command-line interfaces in general.
Instructors likewise may confuse the expected digital literacy of students with computational fluency, even when modern technology increasingly abstracts concepts through search optimization and preponderant IDEs (Integrated Development Environments), or `point-and-click' user interfaces.

Second, while many EEB researchers take advantage of GitHub for individual use, collaborative use may lag due to how researchers traditionally divide labor within projects.
Despite broad utility, GitHub remains a tool predominantly used by computer scientists and software developers.
EEB researchers may take the view that GitHub is a platform that only needs to be used by individuals writing code, and may silo those aspects of projects to a single individual.
Those assumptions may obscure the utility of GitHub for tasks other than traditional data analysis and code development.
However, we emphasize that there are opportunities for collaboration using GitHub by researchers of all skill levels or time constraints (Table \ref{tbl:roles}); for example, project stakeholders can provide a list of use-cases or highlight important conceptual components of a project using GitHub's issues or discussions features.

A third barrier may come from general reluctance to share data and code publicly, or technical and logistical issues\textsuperscript{\protect\hyperlink{ref-pq2Tv1BC}{53}}.
GitHub is, by default, a public and open platform.
This openness may add additional pressure to students and scientists learning to use the platform.
We also note that large file storage is discouraged (and limited) on GitHub but add-ons do exist (e.g.~\url{https://git-lfs.github.com/}) that permit data storage, and increasing integration between platforms (e.g.\textsuperscript{\protect\hyperlink{ref-lx49NGto}{54}}) allows data to be stored away from GitHub and linked dynamically.
We suspect a major additional barrier to EEB researchers is a distinct lack of GitHub help documents for non-English researchers in ecology and evolution, meaning that EEB researchers potentially miss the opportunity to fully understand the importance of version control, reproducibility, and other benefits of GitHub.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{1.0000}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
10 Tips for getting started in GitHub
\end{minipage} \\
\midrule
\endhead
1. \textbf{Check for an existing solution to your problem.} The GitHub Help webpage (\url{https://docs.github.com/en}) contains extensive and detailed documents with helpful screenshots. It is a good starting point for handling an issue, and has troubleshooting tips for specific problems. Alternatively, consider Tweeting your issue. There is a large community of GitHub users around the world who have likely faced analogous problems and may be able to provide quick solutions. Third, try to follow blogs (e.g., \url{https://github.blog}), Twitter accounts or YouTube channels that regularly post practical solutions about the most widely-used web platform for common GitHub issues. \\
2. \textbf{Consider taking free courses.} such as those from Software Carpentry\textsuperscript{\protect\hyperlink{ref-pjy75gHr}{55}} and sharing these courses with your lab members or colleagues. \\
3. \textbf{Take advantage of GitHub as an asynchronous working tool for team-based projects.} See the repository for this paper (\url{https://github.com/SORTEE-Github-Hackathon/manuscript}) as an example of a collaborative manuscript that includes discussions, issues, and a website via GitHub. \\
4. \textbf{The GitHub Learning Lab (\url{https://lab.github.com}).} allows you to learn GitHub basics through short projects and tasks, and allows you to get feedback from their Learning Lab bot. \\
5. \textbf{Check out the following markdown cheatsheet (\url{http://markdownguide.org/basic-syntax}).} so that you can write clear metadata README files for your repositories. \\
6. \textbf{The Jenny Bryan universe of GitHub material.} provides a thorough and accessible introduction for a multitude of research-related uses for GitHub, and includes a book\textsuperscript{\protect\hyperlink{ref-ZvrOcg9w}{56}}, statistics course\textsuperscript{\protect\hyperlink{ref-6CMMeSeD}{57}} and academic article\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}. \\
7. \textbf{Don't be afraid of trial and error.} One of the best ways to learn GitHub is the trial and error method. Learning from your own mistakes can be the better way to master your GitHub abilities. In any case, GitHub has the advantage of making it easy to go back to any steps that you desire via version controlling if you make mistakes. \\
8. \textbf{If you are an educator, include lectures on reproducibility and tools for creating reproducible workflows in the curricula.} Some graduate programs include coursework on course Rmarkdown and GitHub. Getting students started with these tools earlier will prevent the resistance that comes from working with a less reproducible workflow for a longer period of time. (see example \url{https://github.com/rmcelreath/stat_rethinking_2022}) \\
9. \textbf{Try to begin committing with GUI (Graphical user interface) tools.} like GitHub Desktop (\url{https://desktop.github.com}), git-gui (\url{https://git-scm.com/docs/git-gui}), RStudio (\url{https://www.rstudio.com}), Visual Studio Code (\url{https://code.visualstudio.com}), Atom (\url{https://atom.io}), GitKraken (\url{https://www.gitkraken.com}) tools instead of CLI (Command line interface) tools such as Terminal or Console for more advanced features. \\
10. \textbf{Get help deciphering GitHub Notifications.} Try using tools like Octobox (\url{https://octobox.io}) to disentangle and manage multiple notifications from distinct GitHub projects. \\
\bottomrule
\end{longtable}

@@

We provide 12 practical ways that ecologists and evolutionary biologists can incorporate GitHub into their research workflows, and include definitions (\protect\hyperlink{definitions}{Box 1}) and user groups (Figure \ref{fig:github-diagram}) to help researchers prioritize which GitHub skills to learn first.
Some GitHub uses are highly collaborative (\emph{e.g.}, open science discussion and collaborative code editing) while others are focused on individual actions (\emph{e.g.}, storing code/data, building a website).
Regardless of the degree of collaboration, GitHub use in EEB has the potential to make the field more open and transparent than ever before.
Our paper provides the most comprehensive review of how EEB researchers can use GitHub to date, and we encourage researchers at any career stage studying any topic to try GitHub as a platform for sharing and collaboration.

@@

\leavevmode\vadjust pre{\hypertarget{ref-pq2Tv1BC}{}}%
\CSLLeftMargin{53. }
\CSLRightInline{\url{https://doi.org/gqd9cc} doi:\href{https://doi.org/10.31222/osf.io/gaj43}{10.31222/osf.io/gaj43}.}

\leavevmode\vadjust pre{\hypertarget{ref-lx49NGto}{}}%
\CSLLeftMargin{54. }
\CSLRightInline{Connect GitHub to a Project - OSF Support. \url{https://help.osf.io/article/211-connect-github-to-a-project}.}

\leavevmode\vadjust pre{\hypertarget{ref-pjy75gHr}{}}%
\CSLLeftMargin{55. }
\CSLRightInline{Madicken Munk \emph{et al.} swcarpentry/git-novice: Software Carpentry: Version Control with Git, June 2019. (2019) doi:\href{https://doi.org/10.5281/zenodo.3264950}{10.5281/zenodo.3264950}.}

\leavevmode\vadjust pre{\hypertarget{ref-ZvrOcg9w}{}}%
\CSLLeftMargin{56. }
\CSLRightInline{Hester, J. B., the STAT 545 TAs, Jim. \emph{\href{https://happygitwithr.com/}{Let's Git started \textbar{} Happy Git and GitHub for the useR}}.}

\leavevmode\vadjust pre{\hypertarget{ref-6CMMeSeD}{}}%
\CSLLeftMargin{57. }
\CSLRightInline{Bryan, J. \& TAs, T. S. 545. \emph{\href{https://stat545.com/}{STAT 545}}.}
